```{r, setup=TRUE, echo=FALSE}
library(RSQLite)
library(tidyverse)
library(audio)

filename <- "data/wjazzd.db"
sqlite.driver <- dbDriver("SQLite")
db <- dbConnect(sqlite.driver,
                dbname = filename)
dbListTables(db)

tables = list()
for (table in dbListTables(db)) {
  tables[[table]] = dbReadTable(db, table)
}

# Import the midi note table
midiNoteTable = read_csv("data/midiNoteTable.csv")
scaleDegrees = read_csv("data/scaleDegrees.csv")
enharmonicNotes = read_csv("data/enharmonicNotes.csv")
```

The data is sourced from here[https://jazzomat.hfm-weimar.de/download/download.html].

Content[https://jazzomat.hfm-weimar.de/dbformat/dbcontent.html]
Format[https://jazzomat.hfm-weimar.de/dbformat/dbformat.html]
Glossary[https://jazzomat.hfm-weimar.de/dbformat/glossary.html]

## Table info
###beats
Table for beat annotation of WJD melodies, referenced by melody(melid)

###composition_info
Infos regarding the underlying composition of a WJD solo, referenced by melody(melid)

###db_info
Information regarding the distributed database file like version information, license, etc

###esac_info
EsAC infos for EsAC melodies, referenced by melody(melid). 
Note empty because we aren't looking at the EsAC database.

###melody
Main table for all melody events

###melody_type
Indicated type of melody: WJD solos or EsAC (Folk songs using Essen Associative Code), referenced by melody(melid)
Note all SOLO. 

###popsong_info
Pop song infos, referenced by melody(melid)
Note empty

###record_info
Infos regarding the specific audio recording of a WJD solo was taken from, referenced by melody(melid)

###sections
All sections (phrase, chorus, form, chords, etc.), referenced by melody(melid)

###solo_info
Solo infos for WJD solos, referenced by melody(melid)

###track_info
Information specific to a track on a record (or CD)

###transcription_info
Transcription infos for WJD solos, referenced by melody(melid)

# Main tables
THe main tables I care about seem to be:
beats,
composition info,
melody,
record info,
sections,
solo info,
track info,
transcription info

```{r}

getSoloMetaData <- function(melid){
  soloMetadata = tables$solo_info %>%
    filter(melid==!!melid) %>%
    left_join(tables$track_info, by=c("trackid", "recordid", "compid"))
  
  return(soloMetadata)
}

generateMelodyNotes <- function(melid_) {
  soloMelody = tables$melody %>%
    filter(melid==melid_) %>%
    left_join(midiNoteTable, by=c("pitch"="MIDI Note"))
    
  melody = soloMelody %>%
    mutate(Note = paste0(Note, Octave)) %>%
    select(onset, duration, Note)
  
  melodyWithRests = data.frame(onset=as.numeric(), duration=as.numeric(), Note=as.character())
  currentTime = 0
  for (n in 1:nrow(melody)) {
    nextNoteTime = melody$onset[n] - currentTime
    
    if (nextNoteTime > 0) {
      restRow = data.frame(onset=currentTime, duration=nextNoteTime, Note="X")
      melodyWithRests = rbind(melodyWithRests, restRow)
    }
    
    noteRow = melody[n,]
    melodyWithRests = rbind(melodyWithRests, noteRow)
    
    currentTime = melody$onset[n] + melody$duration[n]
     
  }
  
  return (melodyWithRests)
}

generateBassNotes <- function(melid_, bassNoteLength=0.8) {
  soloMetaData = getSoloMetaData(melid_)
  bpm = soloMetaData$avgtempo
  beatLengthSeconds = 1 / (bpm / 60)
  bassNoteLengthSeconds = beatLengthSeconds * bassNoteLength
  restLengthSeconds = beatLengthSeconds * (1-bassNoteLength)
  
  
  beats = tables$beats %>%
    filter(melid == melid_) %>%
    left_join(midiNoteTable, by=c("bass_pitch"="MIDI Note")) %>%
    mutate(Note = paste0(Note, Octave)) %>%
    select(onset, Note)
  
  melodyWithRests = data.frame(onset=as.numeric(), duration=as.numeric(), Note=as.character())

  nextNoteTime = beats$onset[1]
  restRow = data.frame(onset=0, duration=nextNoteTime, Note="X")
  melodyWithRests = rbind(melodyWithRests, restRow)
  
  for (n in 1:nrow(beats)) {
    nextNoteTime = beats$onset[n+1] - beats$onset[n]
    if (is.na(nextNoteTime)) {
      nextNoteTime=1
    }
    noteRow = data.frame(onset=beats$onset[n], duration=bassNoteLengthSeconds, Note=beats$Note[n])
    restRow = data.frame(onset=beats$onset[n] + bassNoteLengthSeconds, duration=restLengthSeconds, Note="X")
    melodyWithRests = rbind(melodyWithRests, noteRow, restRow)
  }
  
  return (melodyWithRests)
}

generateSnareNotes <- function(melid_, snareBeats=c(2, 4), snareLengthSeconds=0.05) {
  beats = tables$beats %>%
    filter(melid == melid_) %>%
    filter(beat %in% snareBeats) %>%
    select(onset)
  
  melodyWithRests = data.frame(onset=as.numeric(), duration=as.numeric(), Note=as.character())

  nextNoteTime = beats$onset[1]
  restRow = data.frame(onset=0, duration=nextNoteTime, Note="X")
  melodyWithRests = rbind(melodyWithRests, restRow)
  
  for (n in 1:nrow(beats)) {
    nextNoteTime = beats$onset[n+1] - beats$onset[n]
    if (is.na(nextNoteTime)) {
      nextNoteTime=1
    }
    noteRow = data.frame(onset=beats$onset[n], duration=snareLengthSeconds, Note="C")
    restRow = data.frame(onset=beats$onset[n]+snareLengthSeconds, duration=nextNoteTime-snareLengthSeconds, Note="X")
    melodyWithRests = rbind(melodyWithRests, noteRow, restRow)
  }
  
  return (melodyWithRests)
}

# generateSnareNotes <- function(melid_, bassNotes, snareBeats=c(2, 4), snareLengthSeconds=0.1) {
#   melodyWithRests = bassNotes %>%
#       left_join(tables$beats %>%
#                   filter(melid == melid_) %>%
#                   select(c("onset", "beat")), by="onset") %>%
#       mutate(Note = ifelse(beat %in% snareBeats, "A", "X")) %>%
#       select(-beat)
#   
#   return (melodyWithRests)
# }

getSoloMelody <- function(melid) {
  soloMetadata = getSoloMetaData(melid)
  
  soloKey = soloMetadata$key
  soloKeyCenter = substr(soloKey, 1, 1)
  
  scaleDegreesLonger = scaleDegrees %>%
    pivot_longer(cols = !Degree, names_to="Key", values_to="Note")
  
  soloScaleDegrees = scaleDegreesLonger %>%
    filter(Key==soloKeyCenter) %>%
    left_join(enharmonicNotes, by="Note")
  
  soloMelody = tables$melody %>%
    filter(melid==!!melid) %>%
    left_join(midiNoteTable, by=c("pitch"="MIDI Note")) %>%
    left_join(soloScaleDegrees %>%
                select(c("Degree", "Enharmonic")), 
              by=c("Note"="Enharmonic")) %>%
    mutate(Degree = fct_relevel(Degree, "1", "b2", "2", "b3", "3", "4", "#4", "5", "b6", "6", "b7", "7"))
  
  return(soloMelody)
}
```

```{r}
sine <- function(freq, duration, note_shortness, tempo, sample_rate) {
  wave <- sin(seq(0, duration/tempo * 60, 1/sample_rate) * 
    freq * 2 * pi)
  
  n = max(2/(length(wave)-2)*sample_rate, note_shortness)
  
  fade <- seq(0, 1, n/sample_rate)
  
  wave * c(fade, rep(1, max(0, length(wave) - 2 * length(fade))), 
  rev(fade))
  
}

whiteNoise <- function(freq, duration, note_shortness, tempo, sample_rate, cutoff=0.02) {
  if (freq == 0) {
    wave <- rep(0, length(seq(0, duration/tempo * 60, 1/sample_rate)))
  }
  else {
    wave <- rnorm(seq(0, duration/tempo * 60, 1/sample_rate))
    wave = wave * (1 / max(wave, na.rm = T))  # normalise
    wave = pmin(cutoff, wave)
    wave = pmax(-cutoff, wave)  # apply filter

  }
  
  n = max(2/(length(wave)-2)*sample_rate, note_shortness)
  
  fade <- seq(0, 1, n/sample_rate)
  
  wave * c(fade, rep(1, max(0, length(wave) - 2 * length(fade))), 
  rev(fade))
}

getWave = function(notes_pitch = NULL, notes_duration = NULL, tempo = 240, note_shortness=100, wave_func=sine) 
{
  . <- NULL
  note <- NULL
  octave <- NULL
  pitch <- NULL
  notes <- c(A = 0, B = 2, C = 3, D = 5, E = 7, F = 8, G = 10)
  
  tune <-
    tibble::data_frame(
      pitch = notes_pitch,
      duration = notes_duration
    ) %>%
    dplyr::mutate(
      octave = substring(pitch, nchar(as.character(pitch))) %>%
        {
          suppressWarnings(as.numeric(.))
        } %>%
        ifelse(is.na(.), 4, .),
      note = ifelse(notes_pitch!="X", 
                    notes[substr(pitch, 1, 1)] + 
                      grepl("#", pitch) -  
                      grepl("b", pitch) + 
                      octave * 12 +
                      12 * (notes[substr(pitch, 1, 1)] < 3),
                    0),
      freq = ifelse(note!=0, 
                    2^((note - 60) / 12) * 440,
                    0
      )
    )
  
  sample_rate <- 44100
 
  tune_wave <- mapply(wave_func, tune$freq, tune$duration, note_shortness, tempo, sample_rate,
    SIMPLIFY = FALSE) %>% do.call("c", .)
}
```

```{r}
getSnareBeats = function(timeSig) {
  if (timeSig == "4/4") {
    return(c(2, 4))
  }
  else if (timeSig == "3/4") {
    return(2)
  }
  else if (timeSig == "6/8") {
    return(4)
  }
  else if (timeSig == "9/4") {
    return(c(2, 5, 8))
  }
  else if (timeSig == "5/4") {
    return(c(2, 4))
  }
  else if (timeSig == "6/4") {
    return(c(3, 6))
  }
  else {
    return(2)
  }
}
```

```{r}
pause(audioInstance)

# melid = 222  # giant steps
# melid = 323  # so what
# melid = 93 # joy spring
# melid = 431  # footprints
melid = 133  # anthropology

soloMetadata = getSoloMetaData(melid)
soloMelodyNotes = generateMelodyNotes(melid)
bassNotes = generateBassNotes(melid)
snareBeats = getSnareBeats(soloMetadata$signature)
snareNotes = generateSnareNotes(melid, snareBeats=snareBeats)

bpm = soloMetadata$avgtem
# snareNotes = snareNotes[1:30, ]

melody = getWave(notes_pitch=soloMelodyNotes$Note, tempo=bpm, notes_duration=soloMelodyNotes$duration*bpm/60, 200)
bass = getWave(notes_pitch=bassNotes$Note, tempo=bpm, notes_duration=bassNotes$duration*bpm/60, 200)
snare = getWave(notes_pitch=snareNotes$Note, tempo=bpm, notes_duration=snareNotes$duration*bpm/60, 200, wave_func=whiteNoise)

addToBass = rep(0, max(0, length(melody)-length(bass), length(snare)-length(bass)))
addToMelody = rep(0, max(0, length(bass)-length(melody), length(snare)-length(melody)))
addToSnare = rep(0, max(0, length(bass)-length(snare), length(melody)-length(snare)))

combined = c(melody, addToMelody) + c(bass, addToBass) + c(snare, addToSnare)
combined = combined * (1 / max(combined, na.rm = T))  # prevent distortion

save.wave(combined, "audio/combined.wav")
wavFile = load.wave("audio/combined.wav")
audioInstance = play(wavFile)
```


